---
title: "Bayesian meta-analysis"
---

```{r}
#| label: setup
#| message: false
#| warning: false
source("analysis/meta_helpers.R")
ma <- load_ma()
bm_primary <- fit_bm(dplyr::filter(ma, tier == "primary"))
bm_sens    <- fit_bm(ma)
```

We fit a Bayesian random-effects (normal-normal) model to the per-study log odds ratios,
with weakly informative priors: a N(0, 1.5) prior on the pooled log-OR and a
half-normal(0.5) prior on the between-study standard deviation τ. The posterior is
computed semi-analytically with `bayesmeta` (an independent MCMC refit and its
convergence diagnostics are on the [Model fit & MCMC](model-fit.qmd) page).

## Pooled estimates

```{r}
#| label: bayes-summary
tibble(
  Analysis = c("Primary (in-hospital / ~30-day)", "Sensitivity (+ longer timepoints)"),
  Studies = c(bm_primary$k, bm_sens$k),
  OR = c(or_ci(bm_primary)[1], or_ci(bm_sens)[1]),
  `CrI lower` = c(or_ci(bm_primary)[2], or_ci(bm_sens)[2]),
  `CrI upper` = c(or_ci(bm_primary)[3], or_ci(bm_sens)[3]),
  `P(OR<1)` = c(bm_primary$pposterior(mu = 0), bm_sens$pposterior(mu = 0))
) |>
  mutate(across(where(is.numeric), ~round(.x, 2))) |>
  knitr::kable()
```

The sensitivity-analysis pooled odds ratio is `{r} fmt(bm_sens)`, with a posterior
probability of benefit (OR < 1) of `{r} sprintf("%.2f", bm_sens$pposterior(mu = 0))`.
The estimate points toward a mortality reduction but the credible interval includes no
effect.

## Posterior forest plot

Each row shows the **posterior distribution** of the study-specific (shrunken) odds
ratio; open points mark the observed estimates. The bottom row is the posterior of the
pooled effect μ.

```{r}
#| label: posterior-forest
set.seed(7183)
ps <- bm_sens$rposterior(n = 6000)   # columns: tau, mu (log-OR scale)

# Conditional posterior of each study's true effect given (mu, tau, y_i, sei_i)
theta_draws <- purrr::map_dfr(seq_len(nrow(ma)), function(i) {
  pd <- 1 / ma$sei[i]^2                 # data precision
  pp <- 1 / ps[, "tau"]^2               # prior (between-study) precision
  v  <- 1 / (pd + pp)
  m  <- v * (ma$yi[i] * pd + ps[, "mu"] * pp)
  tibble(label = ma$study[i], draw = exp(rnorm(nrow(ps), m, sqrt(v))))
})
pooled_draws <- tibble(label = "Pooled (\u03bc)", draw = exp(ps[, "mu"]))
alldraws <- bind_rows(theta_draws, pooled_draws) |>
  mutate(label = factor(label, levels = rev(c(ma$study, "Pooled (\u03bc)"))),
         is_pooled = label == "Pooled (\u03bc)")

obs <- ma |> transmute(label = factor(study, levels = levels(alldraws$label)), OR)

ggplot(alldraws, aes(draw, label)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_violin(aes(fill = is_pooled), orientation = "y", color = NA,
              scale = "width", alpha = 0.85) +
  geom_point(data = obs, aes(OR, label), shape = 21, fill = "white",
             color = jama_blue_grey[["dark"]], size = 2) +
  scale_x_log10(breaks = c(0.1, 0.25, 0.5, 1, 2, 4)) +
  scale_fill_manual(values = c(`FALSE` = jama_blue_grey[["medium"]],
                               `TRUE`  = jama_blue_grey[["accent"]]), guide = "none") +
  labs(x = "Odds ratio (log scale) \u2014 <1 favours combination", y = NULL,
       title = "Posterior forest plot",
       subtitle = "Posterior densities of shrunken study effects and the pooled effect; points = observed ORs") +
  theme_jama()
```

## Marginal posteriors

```{r}
#| label: posterior-mu
mu_grid <- seq(-2.2, 1.6, length.out = 500)
tibble(OR = exp(mu_grid), density = bm_sens$dposterior(mu = mu_grid)) |>
  ggplot(aes(OR, density)) +
  geom_area(fill = jama_blue_grey[["light"]], color = jama_blue_grey[["steel"]]) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_vline(xintercept = or_ci(bm_sens)[1], color = jama_blue_grey[["accent"]]) +
  scale_x_log10(breaks = c(0.1, 0.25, 0.5, 1, 2, 4)) +
  labs(x = "Pooled odds ratio (log scale)", y = "Posterior density",
       title = "Posterior of the pooled effect",
       subtitle = "Solid line = posterior median; dashed = no effect") +
  theme_jama()
```

```{r}
#| label: posterior-tau
tau_grid <- seq(0, 2, length.out = 500)
tibble(tau = tau_grid, density = bm_sens$dposterior(tau = tau_grid)) |>
  ggplot(aes(tau, density)) +
  geom_area(fill = jama_blue_grey[["light"]], color = jama_blue_grey[["steel"]]) +
  labs(x = "Between-study SD (\u03c4, log-OR scale)", y = "Posterior density",
       title = "Posterior of between-study heterogeneity (\u03c4)") +
  theme_jama()
```

The heterogeneity posterior has substantial mass away from zero, consistent with the
moderate I² seen in the [frequentist analysis](frequentist.qmd), but is imprecise given
only six small studies.
