---
title: "Bayesian meta-analysis"
---

```{r}
#| label: setup
#| message: false
#| warning: false
source("analysis/meta_helpers.R")
ma <- load_ma()
bm_primary <- fit_bm(dplyr::filter(ma, tier == "primary"))
bm_sens    <- fit_bm(ma)
```

We fit a Bayesian random-effects (normal-normal) model to the per-study log odds ratios,
with weakly informative priors: a N(0, 1.5) prior on the pooled log-OR and a
half-normal(0.5) prior on the between-study standard deviation τ. The posterior is
computed semi-analytically with `bayesmeta` (an independent MCMC refit and its
convergence diagnostics are on the [Model fit & MCMC](model-fit.qmd) page).

## Pooled estimates

```{r}
#| label: bayes-summary
tibble(
  Analysis = c("Primary (in-hospital / ~30-day)", "Sensitivity (+ longer timepoints)"),
  Studies = c(bm_primary$k, bm_sens$k),
  OR = c(or_ci(bm_primary)[1], or_ci(bm_sens)[1]),
  `CrI lower` = c(or_ci(bm_primary)[2], or_ci(bm_sens)[2]),
  `CrI upper` = c(or_ci(bm_primary)[3], or_ci(bm_sens)[3]),
  `P(OR<1)` = c(bm_primary$pposterior(mu = 0), bm_sens$pposterior(mu = 0))
) |>
  mutate(across(where(is.numeric), ~round(.x, 2))) |>
  knitr::kable()
```

The sensitivity-analysis pooled odds ratio is `{r} fmt(bm_sens)`, with a posterior
probability of benefit (OR < 1) of `{r} sprintf("%.2f", bm_sens$pposterior(mu = 0))`.
The estimate points toward a mortality reduction but the credible interval includes no
effect.

## Posterior forest plot

Each row shows the **posterior distribution** of the study-specific (shrunken) odds
ratio; open points mark the observed estimates. The bottom row is the posterior of the
pooled effect μ.

```{r}
#| label: posterior-forest
set.seed(7183)
ps <- bm_sens$rposterior(n = 6000)   # columns: tau, mu (log-OR scale)

# Conditional posterior of each study's true effect given (mu, tau, y_i, sei_i)
theta_draws <- purrr::map_dfr(seq_len(nrow(ma)), function(i) {
  pd <- 1 / ma$sei[i]^2                 # data precision
  pp <- 1 / ps[, "tau"]^2               # prior (between-study) precision
  v  <- 1 / (pd + pp)
  m  <- v * (ma$yi[i] * pd + ps[, "mu"] * pp)
  tibble(label = ma$study[i], draw = exp(rnorm(nrow(ps), m, sqrt(v))))
})
pooled_draws <- tibble(label = "Pooled (\u03bc)", draw = exp(ps[, "mu"]))
alldraws <- bind_rows(theta_draws, pooled_draws) |>
  mutate(label = factor(label, levels = rev(c(ma$study, "Pooled (\u03bc)"))),
         is_pooled = label == "Pooled (\u03bc)")

obs <- ma |> transmute(label = factor(study, levels = levels(alldraws$label)), OR)

ggplot(alldraws, aes(draw, label)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_violin(aes(fill = is_pooled), orientation = "y", color = NA,
              scale = "width", alpha = 0.85) +
  geom_point(data = obs, aes(OR, label), shape = 21, fill = "white",
             color = jama_blue_grey[["dark"]], size = 2) +
  scale_x_log10(breaks = c(0.1, 0.25, 0.5, 1, 2, 4)) +
  scale_fill_manual(values = c(`FALSE` = jama_blue_grey[["medium"]],
                               `TRUE`  = jama_blue_grey[["accent"]]), guide = "none") +
  labs(x = "Odds ratio (log scale) \u2014 <1 favours combination", y = NULL,
       title = "Posterior forest plot",
       subtitle = "Posterior densities of shrunken study effects and the pooled effect; points = observed ORs") +
  theme_jama()
```

## Marginal posteriors

```{r}
#| label: posterior-mu
mu_grid <- seq(-2.2, 1.6, length.out = 500)
tibble(OR = exp(mu_grid), density = bm_sens$dposterior(mu = mu_grid)) |>
  ggplot(aes(OR, density)) +
  geom_area(fill = jama_blue_grey[["light"]], color = jama_blue_grey[["steel"]]) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_vline(xintercept = or_ci(bm_sens)[1], color = jama_blue_grey[["accent"]]) +
  scale_x_log10(breaks = c(0.1, 0.25, 0.5, 1, 2, 4)) +
  labs(x = "Pooled odds ratio (log scale)", y = "Posterior density",
       title = "Posterior of the pooled effect",
       subtitle = "Solid line = posterior median; dashed = no effect") +
  theme_jama()
```

```{r}
#| label: posterior-tau
tau_grid <- seq(0, 2, length.out = 500)
tibble(tau = tau_grid, density = bm_sens$dposterior(tau = tau_grid)) |>
  ggplot(aes(tau, density)) +
  geom_area(fill = jama_blue_grey[["light"]], color = jama_blue_grey[["steel"]]) +
  labs(x = "Between-study SD (\u03c4, log-OR scale)", y = "Posterior density",
       title = "Posterior of between-study heterogeneity (\u03c4)") +
  theme_jama()
```

The heterogeneity posterior has substantial mass away from zero, consistent with the
moderate I² seen in the [frequentist analysis](frequentist.qmd), but is imprecise given
only six small studies.

## Absolute effect, NNT, and prediction interval

Beyond the odds ratio, two further summaries aid interpretation: the absolute **risk
difference** (with the implied number needed to treat) and a **95% prediction interval** —
the range within which the true effect of a *new* study would be expected to fall. The
prediction interval accounts for between-study heterogeneity and is therefore wider than
the credible interval for the mean effect.

```{r}
#| label: rd-nnt-pi
# Risk difference (Bayesian): mu ~ N(0, 0.5), tau ~ half-normal(0.2) on the RD scale
maRD <- escalc(measure = "RD", ai = e_comb, n1i = n_comb, ci = e_ref, n2i = n_ref,
               data = ma, slab = study)
maRD$sei <- sqrt(maRD$vi)
bm_rd <- bayesmeta(y = maRD$yi, sigma = maRD$sei, labels = maRD$study,
                   mu.prior  = c(mean = 0, sd = 0.5),
                   tau.prior = function(t) dhalfnormal(t, scale = 0.2))
rd <- bm_rd$summary[, "mu"]

# 95% prediction interval (OR scale) from the odds-ratio model's predictive distribution
pred <- exp(bm_sens$qposterior(mu.p = c(0.025, 0.975), predict = TRUE))

tibble(
  Quantity = c("Pooled odds ratio", "Risk difference", "Number needed to treat (point)",
               "95% prediction interval (OR)"),
  Estimate = c(fmt(bm_sens),
               sprintf("%.1f%% (95%% CrI %.1f%% to %.1f%%)",
                       100 * rd["median"], 100 * rd["95% lower"], 100 * rd["95% upper"]),
               sprintf("%.0f (to prevent one death)", 1 / abs(rd["median"])),
               sprintf("%.2f to %.2f", pred[1], pred[2]))
) |> knitr::kable()
```

The pooled risk difference is about `{r} sprintf("%.0f", -100 * rd["median"])` fewer deaths
per 100 patients, but its credible interval spans benefit and harm, so the point NNT of
`{r} sprintf("%.0f", 1 / abs(rd["median"]))` is highly uncertain. The **prediction interval
(OR `{r} sprintf("%.2f-%.2f", pred[1], pred[2])`)** is the key caution: in a new setting the
true effect could plausibly range from a substantial mortality reduction to a doubling of
the odds of death.

## Secondary outcome: clinical response

Three studies report clinical response/cure for both arms. Here the event is a *good*
outcome, so an odds ratio **above 1 favours combination therapy**.

```{r}
#| label: response-pool
resp <- tribble(
  ~study,    ~e_comb, ~n_comb, ~e_ref, ~n_ref, ~confounded,
  "Jin2019",      24,      35,     54,     91, FALSE,
  "Qi2023",       33,      43,     11,     21, FALSE,
  "Li2024a",      20,      20,     12,     18, TRUE)
respE <- escalc(measure = "OR", ai = e_comb, n1i = n_comb, ci = e_ref, n2i = n_ref,
                data = resp, slab = study)
respE$sei <- sqrt(respE$vi); respE$OR <- exp(respE$yi)
bm_resp_clean <- fit_bm(filter(respE, !confounded))
bm_resp_all   <- fit_bm(respE)

respE |>
  transmute(Study = study,
            `Response (comb)` = paste0(e_comb, "/", n_comb),
            `Response (mono)` = paste0(e_ref, "/", n_ref),
            OR = round(OR, 2),
            Note = ifelse(confounded, "confounded; 20/20 \u2192 implausible outlier", "")) |>
  knitr::kable()
```

Excluding the confounded Li2024a (whose 20/20 response yields an implausible OR \u2248 21),
the pooled response odds ratio is `{r} fmt(bm_resp_clean)`; including it, `{r} fmt(bm_resp_all)`.
The direction favours combination therapy and is consistent with Yang et al.'s reported
response benefit (OR 2.30), but our smaller set does not reach conclusiveness (the credible
interval includes 1).

```{r}
#| label: response-forest
est <- respE |>
  transmute(label = study, kind = "study",
            OR = exp(yi), lo = exp(yi - 1.96 * sei), hi = exp(yi + 1.96 * sei))
pool <- tibble(label = c("Pooled (k=2, clean)", "Pooled (k=3, incl. Li2024a)"), kind = "pooled",
  OR = c(or_ci(bm_resp_clean)[1], or_ci(bm_resp_all)[1]),
  lo = c(or_ci(bm_resp_clean)[2], or_ci(bm_resp_all)[2]),
  hi = c(or_ci(bm_resp_clean)[3], or_ci(bm_resp_all)[3]))
bind_rows(est, pool) |>
  mutate(label = factor(label, levels = rev(c("Jin2019", "Qi2023", "Li2024a",
           "Pooled (k=2, clean)", "Pooled (k=3, incl. Li2024a)")))) |>
  ggplot(aes(OR, label, color = kind)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_pointrange(aes(xmin = lo, xmax = hi, shape = kind == "pooled"), linewidth = 0.6) +
  scale_x_log10(breaks = c(0.5, 1, 2, 5, 10, 20)) +
  scale_color_manual(values = c(study = jama_blue_grey[["medium"]],
                                pooled = jama_blue_grey[["accent"]]), guide = "none") +
  scale_shape_manual(values = c(16, 18), guide = "none") +
  labs(x = "Odds ratio (log scale) \u2014 >1 favours combination", y = NULL,
       title = "Clinical response / cure", subtitle = "Event = clinical response") +
  theme_jama()
```

Need for **mechanical ventilation** is reported by several studies, but source
verification found the figures are recorded inconsistently — largely as baseline or
course-of-illness severity (need for ventilation at some point during the ICU stay) rather
than a clearly-timed post-treatment event, and none of the studies time ventilation
relative to treatment start. It is therefore **not pooled** here.
