---
title: "Goodness of fit and model diagnostics"
---

```{r}
#| label: setup
#| message: false
#| warning: false
source("analysis/meta_helpers.R")
ma <- load_ma()
re <- rma(yi, vi, data = ma, method = "REML")
bm_sens <- fit_bm(ma)
```

::: {.callout-note}
Preliminary, machine-assisted diagnostics pending human review. These check whether the
random-effects model is an adequate description of the six source-verified studies.
:::

The [Bayesian](bayesian.qmd) and [frequentist](frequentist.qmd) pages report the pooled
effect; this page asks a different question — **does the model fit the data?** It pairs
frequentist influence diagnostics with Bayesian posterior predictive checks.

## Fit and heterogeneity

```{r}
#| label: fit-summary
ci  <- confint(re)
pri <- predict(re, transf = exp)
tibble(
  Statistic = c("Pooled OR (RE, REML)", "95% CI", "95% prediction interval",
                "tau^2", "I^2 (%)", "H^2", "Q (df)", "Q p-value"),
  Value = c(
    sprintf("%.2f", exp(re$b[1])),
    sprintf("%.2f to %.2f", exp(re$ci.lb), exp(re$ci.ub)),
    sprintf("%.2f to %.2f", pri$pi.lb, pri$pi.ub),
    sprintf("%.3f (95%% CI %.3f to %.3f)", re$tau2,
            ci$random["tau^2", "ci.lb"], ci$random["tau^2", "ci.ub"]),
    sprintf("%.0f (95%% CI %.0f to %.0f)", re$I2,
            ci$random["I^2(%)", "ci.lb"], ci$random["I^2(%)", "ci.ub"]),
    sprintf("%.2f", re$H2),
    sprintf("%.1f (%d)", re$QE, re$k - 1),
    sprintf("%.3f", re$QEp))
) |> knitr::kable()
```

Between-study heterogeneity is moderate (I² ≈ 42%) but imprecise, and Cochran's Q is not
significant. The **prediction interval** is far wider than the confidence interval — a new
study's true effect could plausibly fall on either side of no effect.

## Influence and outlier diagnostics

```{r}
#| label: influence
inf <- influence(re)
tibble(
  Study                 = ma$study,
  `Weight %`            = round(as.numeric(weights(re)), 1),
  `Std. residual`       = round(inf$inf$rstudent, 2),
  hat                   = round(inf$inf$hat, 2),
  DFFITS                = round(inf$inf$dffits, 2),
  `Cook's D`            = round(inf$inf$cook.d, 2),
  `tau^2 (leave-1-out)` = round(inf$inf$tau2.del, 3),
  Influential           = ifelse(inf$inf$inf == "*", "yes", "")
) |> knitr::kable()
```

```{r}
#| label: baujat
tmpf <- tempfile(fileext = ".pdf"); grDevices::pdf(tmpf)
bj <- metafor::baujat(re); grDevices::dev.off(); unlink(tmpf)
bj$study <- ma$study
ggplot(bj, aes(x, y)) +
  geom_point(size = 2.8, color = jama_blue_grey[["steel"]]) +
  geom_text(aes(label = study), size = 3, hjust = -0.15) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.18))) +
  labs(x = "Contribution to overall heterogeneity (Q)",
       y = "Influence on pooled estimate",
       title = "Baujat plot",
       subtitle = "Upper-right = studies driving both heterogeneity and the pooled OR") +
  theme_jama()
```

**Qi2025** is the single influential study (Cook's D ≈ 1.0, standardised residual ≈ +2.3);
removing it collapses τ² to zero, so it is essentially the sole source of heterogeneity and
the main pull toward the null. **Xu2025** is its counterweight, contributing most of the
apparent benefit. This mirrors the [leave-one-out analysis](sensitivity.qmd).

## Posterior predictive checks

We simulate replicated study results from the fitted Bayesian model and compare them with
what was observed. If the model is adequate, observed odds ratios should sit comfortably
within their posterior predictive intervals.

```{r}
#| label: ppc-table
set.seed(5127)
D <- bm_sens$rposterior(n = 10000)   # columns: tau, mu (log-OR scale)
ppc <- purrr::map_dfr(seq_len(nrow(ma)), function(i) {
  theta <- rnorm(nrow(D), D[, "mu"], D[, "tau"])
  yrep  <- rnorm(nrow(D), theta, ma$sei[i])
  pl <- mean(yrep <= ma$yi[i])
  tibble(Study = ma$study[i],
         `Observed OR`   = exp(ma$yi[i]),
         `PP 2.5%`       = exp(unname(quantile(yrep, .025))),
         `PP median`     = exp(unname(quantile(yrep, .5))),
         `PP 97.5%`      = exp(unname(quantile(yrep, .975))),
         `PPP (2-sided)` = 2 * min(pl, 1 - pl))
})

# Omnibus check on Cochran's Q (does the modelled heterogeneity match the data?)
w <- 1 / ma$vi
Qobs <- sum(w * (ma$yi - sum(w * ma$yi) / sum(w))^2)
Qrep <- vapply(seq_len(nrow(D)), function(j) {
  theta <- rnorm(nrow(ma), D[j, "mu"], D[j, "tau"]); yr <- rnorm(nrow(ma), theta, ma$sei)
  mfe <- sum(w * yr) / sum(w); sum(w * (yr - mfe)^2)
}, numeric(1))
q_ppp <- mean(Qrep >= Qobs)

ppc |> mutate(across(where(is.numeric), ~round(.x, 2))) |> knitr::kable()
```

All six observed odds ratios fall within their 95% posterior predictive intervals, and no
two-sided predictive p-value is extreme. The omnibus check on Cochran's Q gives a posterior
predictive p-value of `{r} sprintf("%.2f", q_ppp)`, so the heterogeneity the model expects
is consistent with the data.

```{r}
#| label: ppc-plot
ppc |>
  mutate(Study = factor(Study, levels = rev(ma$study))) |>
  ggplot(aes(y = Study)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_linerange(aes(xmin = `PP 2.5%`, xmax = `PP 97.5%`),
                 color = jama_blue_grey[["light"]], linewidth = 3, alpha = 0.8) +
  geom_point(aes(x = `PP median`), shape = 124, size = 4, color = jama_blue_grey[["steel"]]) +
  geom_point(aes(x = `Observed OR`), shape = 21, fill = "white",
             color = jama_blue_grey[["accent"]], size = 2.8) +
  scale_x_log10(breaks = c(0.1, 0.25, 0.5, 1, 2, 4, 8)) +
  labs(x = "Odds ratio (log scale)", y = NULL,
       title = "Posterior predictive check",
       subtitle = "Bars = 95% posterior predictive interval; white points = observed ORs") +
  theme_jama()
```

## Summary

The random-effects model is an **adequate fit**: no individual study is poorly predicted and
the modelled heterogeneity matches the data. The caveat is not misspecification but
**fragility** — two influential studies (Qi2025 and Xu2025) pulling in opposite directions
account for most of the heterogeneity and the location of the pooled estimate. With only six
small observational studies, the synthesis remains inconclusive.
