---
title: "Model fit and MCMC convergence"
---

```{r}
#| label: setup
#| message: false
#| warning: false
source("analysis/meta_helpers.R")
library(coda)
ma <- load_ma()
```

::: {.callout-note}
Preliminary, machine-assisted output pending human review.
:::

## The model

The headline Bayesian estimates ([Bayesian page](bayesian.qmd)) are computed
semi-analytically by `bayesmeta` and therefore do not involve MCMC. To provide genuine
sampler diagnostics — and an independent check of that computation — we refit the
**identical** model by MCMC using a self-contained random-walk Metropolis sampler
written in base R (no JAGS or Stan dependency).

For studies $$i = 1, \dots, k$$ with observed log odds ratio $$y_i$$ and standard error
$$s_i$$:

$$y_i \sim \mathrm{Normal}(\theta_i,\ s_i^2), \qquad \theta_i \sim \mathrm{Normal}(\mu,\ \tau^2)$$

with weakly informative priors $$\mu \sim \mathrm{Normal}(0, 1.5^2)$$ and
$$\tau \sim \mathrm{half\text{-}Normal}(0.5)$$. The pooled odds ratio is
$$\mathrm{OR} = e^{\mu}$$. The study effects $$\theta_i$$ are integrated out
analytically, so the sampler targets the marginal posterior of $$(\mu, \tau)$$. We ran
4 chains — 3000 burn-in iterations (during which the proposal covariance is adapted)
followed by 10,000 retained iterations per chain.

```{r}
#| label: fit-mcmc
s <- fit_mcmc(ma)
```

## Convergence diagnostics

```{r}
#| label: convergence
rhat <- gelman.diag(s[, c("mu", "tau")], multivariate = FALSE)$psrf
ess  <- effectiveSize(s[, c("mu", "tau", "OR")])

tibble(Parameter = c("mu (log-OR)", "tau"),
       `R-hat` = round(rhat[, 1], 3),
       `R-hat upper CI` = round(rhat[, 2], 3),
       `Effective sample size` = round(ess[c("mu", "tau")])) |>
  knitr::kable()
```

All potential scale reduction factors (R̂) are at or very close to 1.00 and effective
sample sizes are in the thousands, indicating the chains have converged and mixed well.

## Trace and density

```{r}
#| label: trace-plots
draws <- purrr::imap_dfr(s, function(chain, i)
  as_tibble(as.matrix(chain)[, c("mu", "tau")]) |>
    mutate(chain = factor(i), iter = row_number()))
draws_long <- pivot_longer(draws, c(mu, tau), names_to = "parameter", values_to = "value")

ggplot(draws_long, aes(iter, value, color = chain)) +
  geom_line(alpha = 0.6, linewidth = 0.25) +
  facet_wrap(~ parameter, ncol = 1, scales = "free_y") +
  scale_color_manual(values = jama_chains) +
  labs(x = "Iteration", y = NULL, title = "Trace plots (4 chains)") +
  theme_jama()
```

```{r}
#| label: density-plots
ggplot(draws_long, aes(value, color = chain)) +
  geom_density() +
  facet_wrap(~ parameter, ncol = 2, scales = "free") +
  scale_color_manual(values = jama_chains) +
  labs(x = NULL, y = "Density", title = "Posterior density by chain") +
  theme_jama()
```

The chains are indistinguishable, with no drift or stickiness in the traces and
overlapping per-chain densities.

## Cross-check: MCMC vs `bayesmeta`

```{r}
#| label: cross-check
bm_sens  <- fit_bm(ma)
mcmc_or  <- exp(summary(s)$quantiles["mu", c("2.5%", "50%", "97.5%")])
tibble(
  Engine = c("Random-walk Metropolis (MCMC)", "bayesmeta (semi-analytic)"),
  `OR median` = c(mcmc_or[2], or_ci(bm_sens)[1]),
  `CrI lower` = c(mcmc_or[1], or_ci(bm_sens)[2]),
  `CrI upper` = c(mcmc_or[3], or_ci(bm_sens)[3])
) |>
  mutate(across(where(is.numeric), ~round(.x, 3))) |>
  knitr::kable()
```

The two independent implementations agree to within Monte Carlo error, confirming the
semi-analytic posterior used elsewhere on this site.
