---
title: "Risk of bias"
---

```{r}
#| label: setup
#| message: false
#| warning: false
source("analysis/meta_helpers.R")
```

::: {.callout-warning}
**Preliminary assessment.** This is a provisional, machine-assisted mapping to the
ROBINS-I domains, grounded in what was verified during data extraction. It is **not** a
completed formal risk-of-bias assessment. Two reviewers must independently complete
ROBINS-I and reconcile before these judgments are used.
:::

All six included studies are **non-randomised** (observational cohorts or a case series),
so risk of bias is framed with **ROBINS-I**. The dominant concern across the evidence
base is **bias due to confounding**: treatment was assigned by clinicians, not
randomised, so sicker (or, conversely, more salvageable) patients may have
preferentially received the combination regimen.

## Domain-level judgments

```{r}
#| label: rob-data
rob <- tibble::tribble(
  ~study,   ~D1,        ~D2,        ~D3,        ~D4,        ~D5,        ~D6,     ~D7,        ~Overall,
  "Qi2025",  "Serious",  "Moderate", "Low",      "Serious",  "Moderate", "Low",   "Serious",  "Serious",
  "Xu2025",  "Serious",  "Moderate", "Low",      "Serious",  "Moderate", "Low",   "Moderate", "Serious",
  "Lu2017",  "Serious",  "Serious",  "Low",      "Moderate", "Moderate", "Low",   "Moderate", "Serious",
  "Jin2019", "Serious",  "Moderate", "Low",      "Moderate", "Low",      "Low",   "Moderate", "Serious",
  "Qi2023",  "Serious",  "Moderate", "Low",      "Moderate", "Low",      "Low",   "Moderate", "Serious",
  "Li2024a", "Serious",  "Moderate", "Moderate", "Critical", "Low",      "Low",   "Moderate", "Critical"
)

domain_labels <- c(
  D1 = "D1 Confounding", D2 = "D2 Selection of participants",
  D3 = "D3 Classification of interventions", D4 = "D4 Deviations (co-interventions/timing)",
  D5 = "D5 Missing data", D6 = "D6 Measurement of outcome",
  D7 = "D7 Selection of reported result", Overall = "Overall")

rob |> knitr::kable()
```

## Traffic-light plot

```{r}
#| label: rob-plot
levs <- c("Low", "Moderate", "Serious", "Critical")
rob_long <- rob |>
  pivot_longer(-study, names_to = "domain", values_to = "judgment") |>
  mutate(judgment = factor(judgment, levels = levs),
         domain = factor(domain, levels = names(domain_labels), labels = domain_labels),
         study = factor(study, levels = rev(c("Qi2025","Xu2025","Lu2017","Jin2019","Qi2023","Li2024a"))))

# Conventional risk-of-bias traffic-light colours (override the blue-grey palette here).
pal <- c(Low = "#4CAF50", Moderate = "#FBC02D", Serious = "#E53935", Critical = "#B71C1C")
txt <- c(Low = "white", Moderate = "#333333", Serious = "white", Critical = "white")

ggplot(rob_long, aes(domain, study, fill = judgment)) +
  geom_tile(color = "white", linewidth = 1) +
  geom_text(aes(label = substr(judgment, 1, 1), color = judgment),
            size = 3.4, fontface = "bold") +
  scale_fill_manual(values = pal, name = "Concern") +
  scale_color_manual(values = txt, guide = "none") +
  labs(x = NULL, y = NULL, title = "Preliminary ROBINS-I traffic-light plot",
       subtitle = "Green = low, yellow = moderate, red = serious, dark red = critical (L/M/S/C)") +
  theme_jama() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Domain-by-domain assessment (draft for ratification)

Each domain states the ROBINS-I question, the judgement applied, and the study-specific
reasoning drawn from data extraction. **This is a single-pass, machine-assisted draft**: two
reviewers should independently confirm or revise every cell of the table above and then
reconcile (see [Ratification workflow](#ratification-workflow)).

### D1 — Confounding

*Did baseline or time-varying confounding distort the treatment–outcome association?*

**Serious for all six studies.** Treatment was allocated by clinicians, not randomised, so
confounding by indication is the dominant threat. Few studies adjust, and where they do the
results conflict: Jin2019 reports an all-patient adjusted HR ≈ 1.10 (near null), whereas
Xu2025's IPTW/Cox analysis gives HR ≈ 2.97 — opposite in direction to the unadjusted pool,
consistent with its combination arm being sicker at baseline. Qi2025's contrast is an
unadjusted subgroup drawn from a risk-factor study. No study controls a pre-specified
confounder set, so none reaches "moderate".

### D2 — Selection of participants into the study

*Was selection into the study or the analysis related to intervention and outcome?*

**Serious for Lu2017** — a single-centre case series of 5 vs 6 patients, highly susceptible
to selection. **Moderate elsewhere**: retrospective cohorts assembled from records, with
Xu2025 explicitly flagging "selection bias in caspofungin use" and Qi2025 analysing a
subgroup selected from a larger cohort.

### D3 — Classification of interventions

*Were intervention groups clearly defined and recorded at the time of treatment?*

**Low for most** — echinocandin + TMP-SMX vs TMP-SMX are objectively documented from
prescribing records. **Moderate for Li2024a**, whose "combination" arm bundles caspofungin
with a corticosteroid and a lower TMP-SMX dose, blurring what the intervention actually is.

### D4 — Deviations from intended interventions (co-interventions, timing)

*Were there systematic differences in co-interventions or in treatment timing?*

**Critical for Li2024a** — the caspofungin arm's corticosteroid and reduced-TMP co-
interventions are confounded with the antifungal, so an echinocandin effect cannot be
isolated. **Serious for Qi2025 and Xu2025**, where caspofungin timing is mixed
initial/salvage and not cleanly recorded relative to treatment start. **Moderate** for the
remaining studies (Lu2017, Jin2019, Qi2023).

### D5 — Missing data

*Were outcome (and covariate) data reasonably complete?*

**Moderate for the retrospective cohorts Qi2025 and Xu2025 and the tiny Lu2017 series**,
given incomplete covariate/outcome capture and no formal handling of missingness;
**Low for Jin2019, Qi2023 and Li2024a**, where mortality was ascertained for all included
patients.

### D6 — Measurement of the outcome

*Could outcome ascertainment differ between groups?*

**Low for all.** The outcome is all-cause mortality (in-hospital / ~30-day) — an objective
endpoint unlikely to be assessed differentially by treatment arm.

### D7 — Selection of the reported result

*Is the reported effect likely selected from multiple analyses or subgroups?*

**Serious for Qi2025**, whose mortality comparison is a subgroup extracted from a
risk-factor study with no pre-registered analysis plan (the same feature that produced
Yang et al.'s ventilated-subgroup extraction error). **Moderate elsewhere**: retrospective
designs without public protocols, so selective reporting cannot be excluded.

## Overall judgement

Following ROBINS-I (the overall rating tracks the least favourable domain), the body of
evidence is at **serious** risk of bias — driven by confounding — with **Li2024a at
critical** risk from its bundled co-interventions. This is the risk-of-bias input to the
[GRADE assessment](conclusions.qmd#certainty-of-evidence-grade) (serious → −1). It is
unlikely to improve under formal review, because every included study is non-randomised.

## Ratification workflow

To move this from *draft* to *final*:

1. Two reviewers independently rate all 6 studies × 7 domains using the full ROBINS-I
   signalling questions.
2. Record both sets of judgements and reconcile disagreements (a third reviewer breaks ties).
3. Replace the judgements in the table above with the reconciled consensus and remove the
   preliminary banner.

Until then, every judgement here is provisional.
