---
title: "Risk of bias"
---

```{r}
#| label: setup
#| message: false
#| warning: false
source("analysis/meta_helpers.R")
```

::: {.callout-warning}
**Preliminary assessment.** This is a provisional, machine-assisted mapping to the
ROBINS-I domains, grounded in what was verified during data extraction. It is **not** a
completed formal risk-of-bias assessment. Two reviewers must independently complete
ROBINS-I and reconcile before these judgments are used.
:::

All six included studies are **non-randomised** (observational cohorts or a case series),
so risk of bias is framed with **ROBINS-I**. The dominant concern across the evidence
base is **bias due to confounding**: treatment was assigned by clinicians, not
randomised, so sicker (or, conversely, more salvageable) patients may have
preferentially received the combination regimen.

## Domain-level judgments

```{r}
#| label: rob-data
rob <- tibble::tribble(
  ~study,   ~D1,        ~D2,        ~D3,        ~D4,        ~D5,        ~D6,     ~D7,        ~Overall,
  "Qi2025",  "Serious",  "Moderate", "Low",      "Serious",  "Moderate", "Low",   "Serious",  "Serious",
  "Xu2025",  "Serious",  "Moderate", "Low",      "Serious",  "Moderate", "Low",   "Moderate", "Serious",
  "Lu2017",  "Serious",  "Serious",  "Low",      "Moderate", "Moderate", "Low",   "Moderate", "Serious",
  "Jin2019", "Serious",  "Moderate", "Low",      "Moderate", "Low",      "Low",   "Moderate", "Serious",
  "Qi2023",  "Serious",  "Moderate", "Low",      "Moderate", "Low",      "Low",   "Moderate", "Serious",
  "Li2024a", "Serious",  "Moderate", "Moderate", "Critical", "Low",      "Low",   "Moderate", "Critical"
)

domain_labels <- c(
  D1 = "D1 Confounding", D2 = "D2 Selection of participants",
  D3 = "D3 Classification of interventions", D4 = "D4 Deviations (co-interventions/timing)",
  D5 = "D5 Missing data", D6 = "D6 Measurement of outcome",
  D7 = "D7 Selection of reported result", Overall = "Overall")

rob |> knitr::kable()
```

## Traffic-light plot

```{r}
#| label: rob-plot
levs <- c("Low", "Moderate", "Serious", "Critical")
rob_long <- rob |>
  pivot_longer(-study, names_to = "domain", values_to = "judgment") |>
  mutate(judgment = factor(judgment, levels = levs),
         domain = factor(domain, levels = names(domain_labels), labels = domain_labels),
         study = factor(study, levels = rev(c("Qi2025","Xu2025","Lu2017","Jin2019","Qi2023","Li2024a"))))

pal <- c(Low = jama_blue_grey[["light"]], Moderate = jama_blue_grey[["medium"]],
         Serious = jama_blue_grey[["steel"]], Critical = jama_blue_grey[["accent"]])

ggplot(rob_long, aes(domain, study, fill = judgment)) +
  geom_tile(color = "white", linewidth = 1) +
  geom_text(aes(label = substr(judgment, 1, 1),
                color = judgment %in% c("Serious", "Critical")), size = 3.4, fontface = "bold") +
  scale_fill_manual(values = pal, name = "Concern") +
  scale_color_manual(values = c(`TRUE` = "white", `FALSE` = "#243B53"), guide = "none") +
  labs(x = NULL, y = NULL, title = "Preliminary ROBINS-I traffic-light plot",
       subtitle = "Darker = higher concern; L/M/S/C = Low/Moderate/Serious/Critical") +
  theme_jama() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Rationale

- **D1 Confounding — Serious for all.** Non-randomised assignment; confounding by
  indication is likely. Xu2025's monotherapy arm was systematically sicker; Qi2025's
  comparison is an unadjusted subgroup from a risk-factor study.
- **D4 Deviations / co-interventions — Critical for Li2024a.** Its combination arm bundles
  caspofungin with a corticosteroid and a lower TMP dose, so an echinocandin effect cannot
  be isolated. Mixed initial/salvage timing raises concern for Qi2025 and Xu2025.
- **D2 Selection — Serious for Lu2017**, a tiny single-centre case series (n = 5 vs 6).
- **D6 Outcome measurement — Low for all**: all-cause mortality is an objective endpoint.

Overall, the body of evidence is at **serious** risk of bias (driven by confounding),
with Li2024a at **critical** risk. This reinforces the cautious, inconclusive
interpretation of the pooled estimate.
