---
title: "Sensitivity analyses"
---

```{r}
#| label: setup
#| message: false
#| warning: false
source("analysis/meta_helpers.R")
ma <- load_ma()
bm_sens <- fit_bm(ma)
```

## Leave-one-out

Each row omits one study and refits the 6-study model.

```{r}
#| label: loo
loo <- purrr::map_dfr(ma$study, function(s_) {
  bm <- fit_bm(dplyr::filter(ma, study != s_))
  tibble(removed = paste0("\u2212 ", s_),
         OR = or_ci(bm)[1], lo = or_ci(bm)[2], hi = or_ci(bm)[3])
})
loo <- bind_rows(
  tibble(removed = "Full model (k=6)",
         OR = or_ci(bm_sens)[1], lo = or_ci(bm_sens)[2], hi = or_ci(bm_sens)[3]),
  loo)
loo |> mutate(across(where(is.numeric), ~round(.x, 2))) |> knitr::kable()
```

```{r}
#| label: loo-plot
loo |>
  mutate(kind = ifelse(removed == "Full model (k=6)", "full", "loo"),
         removed = factor(removed, levels = rev(removed))) |>
  ggplot(aes(OR, removed, color = kind)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_pointrange(aes(xmin = lo, xmax = hi), shape = 18, linewidth = 0.6) +
  scale_x_log10(breaks = c(0.25, 0.5, 1, 2)) +
  scale_color_manual(values = c(full = jama_blue_grey[["accent"]],
                                loo = jama_blue_grey[["medium"]]), guide = "none") +
  labs(x = "Pooled odds ratio (log scale) \u2014 <1 favours combination", y = NULL,
       title = "Leave-one-out sensitivity", subtitle = "Top row = full model") +
  theme_jama()
```

The pooled OR ranges from about 0.59 (omitting Qi2025) to 0.94 (omitting Xu2025). Every
leave-one-out estimate still crosses OR = 1, so no single study changes the qualitative
conclusion, but **Xu2025 is the single most influential study**, contributing most of the
apparent benefit.

## Prior sensitivity

```{r}
#| label: prior-sensitivity
fit_prior <- function(mu_sd, tau_scale, fam) {
  tp <- switch(fam,
    HN = function(t) dhalfnormal(t, scale = tau_scale),
    HC = function(t) dhalfcauchy(t, scale = tau_scale),
    U  = function(t) dunif(t, 0, tau_scale))
  bayesmeta(y = ma$yi, sigma = ma$sei, labels = ma$study,
            mu.prior = c(mean = 0, sd = mu_sd), tau.prior = tp)
}
grid <- tibble::tribble(
  ~prior,                           ~mu_sd, ~tau_scale, ~fam,
  "Base: mu~N(0,1.5), tau~HN(0.5)",    1.5,  0.50, "HN",
  "Tighter tau ~ HN(0.25)",            1.5,  0.25, "HN",
  "Wider tau ~ HN(1.0)",               1.5,  1.00, "HN",
  "tau ~ half-Cauchy(0.5)",            1.5,  0.50, "HC",
  "tau ~ Uniform(0,2)",                1.5,  2.00, "U",
  "Wider mu ~ N(0,10)",                 10,  0.50, "HN",
  "Tighter mu ~ N(0,1)",               1.0,  0.50, "HN")

purrr::pmap_dfr(list(grid$prior, grid$mu_sd, grid$tau_scale, grid$fam),
  function(p, m, ts, f) {
    bm <- fit_prior(m, ts, f)
    tibble(Prior = p, OR = or_ci(bm)[1], `CrI lower` = or_ci(bm)[2],
           `CrI upper` = or_ci(bm)[3], tau = bm$summary["median", "tau"])
  }) |>
  mutate(across(where(is.numeric), ~round(.x, 2))) |>
  knitr::kable()
```

The pooled OR is essentially invariant to the prior specification (0.75–0.77). Only the
credible-interval width responds to the heterogeneity prior, as expected. The pooled
location is therefore robust; residual uncertainty is driven by the small number of
studies and genuine heterogeneity, not by prior choice.

## Timing subgroups

Classifying studies by treatment timing (following Yang et al. 2026):

```{r}
#| label: timing
bm_initial <- fit_bm(dplyr::filter(ma, timing == "initial"))
bm_mixed   <- fit_bm(dplyr::filter(ma, timing == "mixed"))
tibble(
  Subgroup = c("Initial (Jin2019, Qi2023, Li2024a)", "Mixed (Qi2025, Xu2025, Lu2017)"),
  Studies = c(bm_initial$k, bm_mixed$k),
  OR = c(or_ci(bm_initial)[1], or_ci(bm_mixed)[1]),
  `CrI lower` = c(or_ci(bm_initial)[2], or_ci(bm_mixed)[2]),
  `CrI upper` = c(or_ci(bm_initial)[3], or_ci(bm_mixed)[3])
) |>
  mutate(across(where(is.numeric), ~round(.x, 2))) |>
  knitr::kable()
```

No pure salvage-vs-monotherapy study is available in our verified set, so a salvage
subgroup is not estimable. Both subgroups remain inconclusive.

### Formal timing test and comparison with Yang et al. 2026

```{r}
#| label: timing-vs-yang
re_over <- rma(yi, vi, data = ma, method = "REML")
re_init <- rma(yi, vi, data = dplyr::filter(ma, timing == "initial"), method = "REML")
re_mix  <- rma(yi, vi, data = dplyr::filter(ma, timing == "mixed"),   method = "REML")
re_mod  <- rma(yi, vi, mods = ~factor(timing), data = ma, method = "REML")

ours <- tibble(Analysis = c("Overall (k=6)", "Initial (k=3)", "Mixed (k=3)"),
  OR = c(exp(re_over$b), exp(re_init$b), exp(re_mix$b)),
  lo = c(exp(re_over$ci.lb), exp(re_init$ci.lb), exp(re_mix$ci.lb)),
  hi = c(exp(re_over$ci.ub), exp(re_init$ci.ub), exp(re_mix$ci.ub)),
  source = "Ours (REML)")
yang <- tibble(Analysis = c("Overall", "Initial"), OR = c(0.93, 0.50),
  lo = c(0.71, 0.32), hi = c(1.21, 0.77), source = "Yang 2026")
bind_rows(ours, yang) |> mutate(across(c(OR, lo, hi), ~round(.x, 2))) |> knitr::kable()
```

The **within-data subgroup (moderator) test finds no timing effect modification**
(Q~M~(1) = `{r} sprintf("%.2f", re_mod$QM)`, p = `{r} sprintf("%.3f", re_mod$QMp)`): our
initial and mixed subgroups are nearly identical (OR ≈ 0.70 vs 0.74). This differs from
Yang et al. 2026, whose protective signal is concentrated in the *initial-therapy* subgroup
(OR 0.50) against a null overall (0.93).

```{r}
#| label: timing-forest
comp <- bind_rows(ours, yang) |> mutate(label = paste0(Analysis, " \u2014 ", source))
comp$label <- factor(comp$label, levels = rev(comp$label))
ggplot(comp, aes(OR, label, color = source)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_pointrange(aes(xmin = lo, xmax = hi), linewidth = 0.6, shape = 18) +
  scale_x_log10(breaks = c(0.25, 0.5, 0.75, 1, 1.5)) +
  scale_color_manual(values = c(`Ours (REML)` = jama_blue_grey[["dark"]],
                                `Yang 2026` = jama_blue_grey[["medium"]])) +
  labs(x = "Mortality odds ratio (log scale) \u2014 <1 favours combination", y = NULL,
       title = "Mortality by treatment timing: ours vs Yang et al. 2026") +
  theme_jama()
```

Two caveats. The subgroup test has very low power (k = 3 per arm), so a non-significant
result reflects near-identical point estimates rather than proof of no difference, and our
"initial" tier contains the confounded Li2024a. A direct significance test of our estimates
against Yang's is not valid, since the two analyses share roughly nine studies — the
comparison above is descriptive. The earlier "timing explains the gap" reading held only
under a different regrouping (dropping the mixed Qi2025 → OR 0.59), not from timing acting as
an effect modifier within our verified set.
