---
title: "Sensitivity analyses"
---

```{r}
#| label: setup
#| message: false
#| warning: false
source("analysis/meta_helpers.R")
ma <- load_ma()
bm_sens <- fit_bm(ma)
```

## Leave-one-out

Each row omits one study and refits the 6-study model.

```{r}
#| label: loo
loo <- purrr::map_dfr(ma$study, function(s_) {
  bm <- fit_bm(dplyr::filter(ma, study != s_))
  tibble(removed = paste0("\u2212 ", s_),
         OR = or_ci(bm)[1], lo = or_ci(bm)[2], hi = or_ci(bm)[3])
})
loo <- bind_rows(
  tibble(removed = "Full model (k=6)",
         OR = or_ci(bm_sens)[1], lo = or_ci(bm_sens)[2], hi = or_ci(bm_sens)[3]),
  loo)
loo |> mutate(across(where(is.numeric), ~round(.x, 2))) |> knitr::kable()
```

```{r}
#| label: loo-plot
loo |>
  mutate(kind = ifelse(removed == "Full model (k=6)", "full", "loo"),
         removed = factor(removed, levels = rev(removed))) |>
  ggplot(aes(OR, removed, color = kind)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_pointrange(aes(xmin = lo, xmax = hi), shape = 18, linewidth = 0.6) +
  scale_x_log10(breaks = c(0.25, 0.5, 1, 2)) +
  scale_color_manual(values = c(full = jama_blue_grey[["accent"]],
                                loo = jama_blue_grey[["medium"]]), guide = "none") +
  labs(x = "Pooled odds ratio (log scale) \u2014 <1 favours combination", y = NULL,
       title = "Leave-one-out sensitivity", subtitle = "Top row = full model") +
  theme_jama()
```

The pooled OR ranges from about 0.59 (omitting Qi2025) to 0.94 (omitting Xu2025). Every
leave-one-out estimate still crosses OR = 1, so no single study changes the qualitative
conclusion, but **Xu2025 is the single most influential study**, contributing most of the
apparent benefit.

## Prior sensitivity

```{r}
#| label: prior-sensitivity
fit_prior <- function(mu_sd, tau_scale, fam) {
  tp <- switch(fam,
    HN = function(t) dhalfnormal(t, scale = tau_scale),
    HC = function(t) dhalfcauchy(t, scale = tau_scale),
    U  = function(t) dunif(t, 0, tau_scale))
  bayesmeta(y = ma$yi, sigma = ma$sei, labels = ma$study,
            mu.prior = c(mean = 0, sd = mu_sd), tau.prior = tp)
}
grid <- tibble::tribble(
  ~prior,                           ~mu_sd, ~tau_scale, ~fam,
  "Base: mu~N(0,1.5), tau~HN(0.5)",    1.5,  0.50, "HN",
  "Tighter tau ~ HN(0.25)",            1.5,  0.25, "HN",
  "Wider tau ~ HN(1.0)",               1.5,  1.00, "HN",
  "tau ~ half-Cauchy(0.5)",            1.5,  0.50, "HC",
  "tau ~ Uniform(0,2)",                1.5,  2.00, "U",
  "Wider mu ~ N(0,10)",                 10,  0.50, "HN",
  "Tighter mu ~ N(0,1)",               1.0,  0.50, "HN")

purrr::pmap_dfr(list(grid$prior, grid$mu_sd, grid$tau_scale, grid$fam),
  function(p, m, ts, f) {
    bm <- fit_prior(m, ts, f)
    tibble(Prior = p, OR = or_ci(bm)[1], `CrI lower` = or_ci(bm)[2],
           `CrI upper` = or_ci(bm)[3], tau = bm$summary["median", "tau"])
  }) |>
  mutate(across(where(is.numeric), ~round(.x, 2))) |>
  knitr::kable()
```

The pooled OR is essentially invariant to the prior specification (0.75–0.77). Only the
credible-interval width responds to the heterogeneity prior, as expected. The pooled
location is therefore robust; residual uncertainty is driven by the small number of
studies and genuine heterogeneity, not by prior choice.

## Timing subgroups

Classifying studies by treatment timing (following Yang et al. 2026):

```{r}
#| label: timing
bm_initial <- fit_bm(dplyr::filter(ma, timing == "initial"))
bm_mixed   <- fit_bm(dplyr::filter(ma, timing == "mixed"))
tibble(
  Subgroup = c("Initial (Jin2019, Qi2023, Li2024a)", "Mixed (Qi2025, Xu2025, Lu2017)"),
  Studies = c(bm_initial$k, bm_mixed$k),
  OR = c(or_ci(bm_initial)[1], or_ci(bm_mixed)[1]),
  `CrI lower` = c(or_ci(bm_initial)[2], or_ci(bm_mixed)[2]),
  `CrI upper` = c(or_ci(bm_initial)[3], or_ci(bm_mixed)[3])
) |>
  mutate(across(where(is.numeric), ~round(.x, 2))) |>
  knitr::kable()
```

No pure salvage-vs-monotherapy study is available in our verified set, so a salvage
subgroup is not estimable. Both subgroups remain inconclusive.
