---
title: "Included studies"
---

```{r}
#| label: setup
#| message: false
#| warning: false
source("analysis/meta_helpers.R")
ma <- load_ma()
```

## Analytic dataset

All six included studies are **observational** (retrospective cohorts, or a small case
series for Lu2017) of HIV-negative / immunocompromised adults. Each contributes one
two-arm comparison: adjunctive echinocandin/caspofungin + TMP-SMX versus TMP-SMX
monotherapy, with all-cause mortality at the timepoint shown.

```{r}
#| label: studies-table
ma |>
  transmute(Study = study, Tier = tier, Timing = timing,
            `Mortality timepoint` = timepoint,
            `Combination (deaths/N)` = paste0(e_comb, "/", n_comb),
            `Monotherapy (deaths/N)` = paste0(e_ref, "/", n_ref),
            OR = round(OR, 2),
            `95% CI` = sprintf("%.2f\u2013%.2f", exp(yi - 1.96 * sei), exp(yi + 1.96 * sei))) |>
  knitr::kable()
```

The **primary** tier comprises studies reporting in-hospital or ~30-day mortality; the
**sensitivity** tier adds studies reporting longer-timepoint mortality. Treatment
**timing** (initial vs mixed) is classified following Yang et al. (2026); no
salvage-vs-monotherapy study is available.

## Verification notes

Every count above was checked against the source full text. Key decisions:

| Study | Note |
|---|---|
| Qi2025 | Full-cohort Table 4 (18/35 vs 31/79). The same PUMCH cohort as Yanmeng2026 (duplicate removed). Prior meta-analyses that report 18/27 used the *ventilated subgroup*. |
| Xu2025 | Multicentre cohort; the monotherapy arm was sicker (more solid tumour, more aspergillosis), so the crude comparison is subject to confounding by indication. Most influential study (see [sensitivity](sensitivity.qmd)). |
| Lu2017 | Very small case series (n = 5 vs 6), heart-transplant recipients; mortality timepoint not explicitly defined. |
| Jin2019 | 3-month mortality; full-cohort arms (not the β-D-glucan subgroups). |
| Qi2023 | 90-day mortality; the *initial* caspofungin arm used (salvage arm excluded from this contrast). |
| Li2024a | **Confounded**: the combination arm bundles caspofungin with a corticosteroid and a lower TMP dose, so it cannot isolate an echinocandin effect. Retained in the sensitivity tier only, flagged. |

## Studies not analysed

- **Tian2020** — excluded: an HIV-*positive* cohort (published in *HIV Medicine*), outside the population.
- **Wang2019 pilot, Zhang2018** — no TMP-SMX monotherapy comparator arm.
- **Taniguchi (administrative dataset)** — mortality events were back-calculated; excluded from the primary synthesis.
- **Seven studies in Yang et al. that we could not retrieve** — mostly Chinese-language papers / theses (CNKI/Wanfang); one (Wang-ZG-2019) was obtained and found ineligible (single-arm). See [conclusions](conclusions.qmd).
